\vspace{1.0cm} \LARGE \bf Reassortment between influenza B lineages and the emergence of a co-adapted PB1-PB2-HA gene complex

### 2014-05-15 fluB.tex
Add extra acknowledgments.
Darren now owes me a beer :D
Clarify null probability calculation per #12.

We reconstructed reassortment events that were detected by using lineage labels.
Figure \ref{railroadPlot} focuses only on inter-lineage reassortments that have occurred after 1990.
We identify 5 major (in terms of persistence) reassortant genome constellations (given in order PB1-PB2-PA-HA-NP-NA-MP-NS with prime (') indicating independently acquired segments) circulating between 1992 and 2011 (Figure \ref{railroadPlot}):
\begin{itemize}
  \item B/Alaska/12/1996-like (Y-Y-Y-Y-Y-Y-Y-V)
  \item B/Nanchang/2/1997-like (V-V-Y-V-Y-V-Y-V)
  \item B/Iowa/03/2002-like (V-V-Y'-V-Y-Y-Y'-V')
  \item B/California/NHRC0001/2006-like (V-V-Y-V-Y'-Y-Y'-V')
  \item B/Brisbane/33/2008-like (V-V-Y-V-Y'-Y-Y-V)
\end{itemize}
B/Alaska/12/1996-like, B/Nanchang/2/1997-like and B/Iowa/03/2002-like constellations were observed by Chen \& Holmes \citep{chen2008}, but sequences from \\B/California/NHRC0001/2006-like and B/Brisbane/33/2008-like constellations were not available at the time.
Of these 5 constellations 4 (B/Nanchang/2/1997-like, B/Iowa/03/2002-like, B/California/NHRC0001/2006-like and B/Brisbane/33/2008-like) are derived from introgression of Yamagata lineage segments into Victoria lineage PB1-PB2-HA background, with only 1 (B/Alaska/12/1996-like) resulting from introgression of Victoria lineage NS segment into an entirely Yamagata derived background.
All 5 inter-lineage reassortment events described here are marked by the preservation of either entirely Victoria or Yamagata derived PB1-PB2-HA segments.
Figure \ref{railroadPlot} also shows that reassorting segments appear to evolve with a considerable degree of autonomy.
For example, the NP lineage that entered a largely Victoria lineage derived genome and gave rise to the B/Nanchang/2/1997-like isolates continued circulating until 2010, even though other segments it reassorted with in 1995 -- 1996 (PA and MP) went extinct following the reassortments that led to the rise of viruses with B/Iowa/03/2002-like genome constellations.
A more extreme example is the NS segment, which in B/Iowa/03/2002-like isolates (and all subsequent Vic PB1-PB2-HA isolates) has originally been derived from the Victoria lineage that had been associated with mostly Yam lineage derived B/Alaska/12/1996-like genomes for a number of years.
Overall, we see that successful inter-lineage reassortment events do not break up Vic and Yam PB1-PB2-HA complexes, but any other segments reassorted into one or the other PB1-PB2-HA background evolve with a considerable degree of autonomy.

@@

\section*{Acknowledgements}
GD was supported by a Natural Environment Research Council studentship D76739X.
TB was supported by a Newton International Fellowship from the Royal Society. 
The research leading to these results has received funding from the European Research Council under the European Community's Seventh Framework Programme (FP7/2007-2013) under Grant Agreement no. 278433-PREDEMICS and ERC Grant agreement no. 260864.
AR and SL acknowledge the support of the Wellcome Trust (grant no. 092807).

>>>>

We reconstructed reassortment events that were detected by using lineage labels.
Figure \ref{railroadPlot} focuses only on inter-lineage reassortments that have occurred after 1990.
We identify 5 major (in terms of persistence) reassortant genome constellations (given in order PB1-PB2-PA-HA-NP-NA-MP-NS with prime (') indicating independently acquired segments) circulating between 1992 and 2011 (Figure \ref{railroadPlot}):
\begin{itemize}
  \item B/Alaska/12/1996-like (Y-Y-Y-Y-Y-Y-Y-V)
  \item B/Nanchang/2/1997-like (V-V-Y-V-Y-V-Y-V)
  \item B/Iowa/03/2002-like (V-V-Y'-V-Y-Y-Y'-V')
  \item B/California/NHRC0001/2006-like (V-V-Y-V-Y'-Y-Y'-V')
  \item B/Brisbane/33/2008-like (V-V-Y-V-Y'-Y-Y-V)
\end{itemize}
B/Alaska/12/1996-like, B/Nanchang/2/1997-like and B/Iowa/03/2002-like constellations were observed by Chen \& Holmes \citep{chen2008}, but sequences from \\B/California/NHRC0001/2006-like and B/Brisbane/33/2008-like constellations were not available at the time.
Of these 5 constellations 4 (B/Nanchang/2/1997-like, B/Iowa/03/2002-like, B/California/NHRC0001/2006-like and B/Brisbane/33/2008-like) are derived from introgression of Yamagata lineage segments into Victoria lineage PB1-PB2-HA background, with only 1 (B/Alaska/12/1996-like) resulting from introgression of Victoria lineage NS segment into an entirely Yamagata derived background.
All 5 inter-lineage reassortment events described here are marked by the preservation of either entirely Victoria or Yamagata derived PB1-PB2-HA segments.
Figure \ref{railroadPlot} also shows that reassorting segments appear to evolve with a considerable degree of autonomy.
For example, the NP lineage that entered a largely Victoria lineage derived genome and gave rise to the B/Nanchang/2/1997-like isolates continued circulating until 2010, even though other segments it reassorted with in 1995 -- 1996 (PA and MP) went extinct following the reassortments that led to the rise of viruses with B/Iowa/03/2002-like genome constellations.
A more extreme example is the NS segment, which in B/Iowa/03/2002-like isolates (and all subsequent Vic PB1-PB2-HA isolates) has originally been derived from the Victoria lineage that had been associated with mostly Yam lineage derived B/Alaska/12/1996-like genomes for a number of years.
We observe that in 5 successful inter-lineage reassortment events, none break up the PB1-PB2-HA complex.
This is an unlikely outcome -- the probability of not breaking up PB1-PB2-HA across 5 reassortment events is $p = (\frac{2^{5} \times 2 - 2}{2^{8} - 2})^{5} = 0.0009$, where reassortment events are considered to sample from the Vic and Yam lineages at random for each of the 8 segments.
If we correct for multiple testing with the assumption that co-assortment of any 3 segments is of interest we find that the probability of not breaking up an arbitrary 3 segments across 5 reassortment events is $p = \binom{8}{3} \times (\frac{2^{5} \times 2 - 2}{2^{8} - 2})^{5} = 0.0485$.

@@

\section*{Acknowledgements}
We would like to thank Darren Obbard and Paul Wikramaratna for helpful discussions.
GD was supported by a Natural Environment Research Council studentship D76739X.
TB was supported by a Newton International Fellowship from the Royal Society. 
The research leading to these results has received funding from the European Research Council under the European Community's Seventh Framework Programme (FP7/2007-2013) under Grant Agreement no. 278433-PREDEMICS and ERC Grant agreement no. 260864.
AR and SL acknowledge the support of the Wellcome Trust (grant no. 092807).
